How Short Lived is the Population III Epoch?

### 2010-08-25 ms.tex
Started the method section.

\author{John H. Wise\altaffilmark{1,2}, 
  Michael L. Norman\altaffilmark{3},
  Tom Abel\altaffilmark{4}}

\altaffiltext{1}{Department of Astrophysical Sciences, Princeton
  University, Peyton Hall, Ivy Lane, Princeton, NJ 08544}
\altaffiltext{2}{Hubble Fellow}
\altaffiltext{3}{Center for Astrophysics and Space Sciences,
  University of California at San Diego, La Jolla, CA 92093}
\altaffiltext{4}{Kavli Institute for Particle Astrophysics and
  Cosmology, Stanford University, Menlo Park, CA 94025}
\email{jwise@astro.princeton.edu}

@@

Numerical simulations are useful to detangle and study these
complexities and the transition from Pop III to II stars.  In this
Letter, we present the simulations that include both types of star
formation, its radiative feedback, and its SNe feedback.  We utilize
an initial mass function (IMF) for individual Pop III stars, based on
the latest simulation results.  The methods used here incorporate and
link together recent results from metal-enriched and metal-free star
formation, the critical metallicity, pair-instability supernovae, and
fine-structure metal line cooling.  Furthermore we spatially
distinguish metal enrichment from Pop II and Pop III stars as their
abundance patterns differ significantly \citep{ref}.  This is the
first time that all of these processes have been interconnected in a
simulation that resolves minihalos to study the transition to Pop II
star formation self-consistently.  We describe our simulation setup
and our adopted star formation models in \S \ref{sec:setup} and
present our results in \S \ref{sec:results}.  We discuss the
implications of our findings and conclude in the last section.

\section{Simulation Setup}
\label{sec:setup}

>>>>

\author{John H. Wise\altaffilmark{1,2}, 
  Michael L. Norman\altaffilmark{3},
  Tom Abel\altaffilmark{4},
  Britton D. Smith \altaffilmark{5}}

\altaffiltext{1}{Department of Astrophysical Sciences, Princeton
  University, Peyton Hall, Ivy Lane, Princeton, NJ 08544}
\altaffiltext{2}{Hubble Fellow}
\altaffiltext{3}{Center for Astrophysics and Space Sciences,
  University of California at San Diego, La Jolla, CA 92093}
\altaffiltext{4}{Kavli Institute for Particle Astrophysics and
  Cosmology, Stanford University, Menlo Park, CA 94025}
\altaffiltext{5}{Michigan State University}
\email{jwise@astro.princeton.edu}

@@

Numerical simulations are useful to detangle and study these
complexities and the transition from Pop III to II stars.  In this
Letter, we present the simulations that include both types of star
formation, its radiative feedback, and its SNe feedback.  We utilize
an initial mass function (IMF) for individual Pop III stars, based on
the latest simulation results.  The methods used here incorporate and
link together recent results from metal-enriched and metal-free star
formation, the critical metallicity, pair-instability supernovae, and
fine-structure metal line cooling.  Furthermore we spatially
distinguish metal enrichment from Pop II and Pop III stars.  This is
the first time that all of these processes have been interconnected in
a simulation that resolves minihalos to study the transition to Pop II
star formation self-consistently.  We describe our simulation setup
and our adopted star formation models in \S \ref{sec:setup} and
present our results in \S \ref{sec:results}.  We discuss the
implications of our findings and conclude in the last section.

\section{Method}
\label{sec:setup}

In this section, we first describe our simulation setup.  We then
detail the star formation models used in these simulations and how we
model stellar radiation and SNe feedback.

\subsection{Simulation setup}

We use the adaptive mesh refinement (AMR) code \enzo~\citep{refs} that
employs our implementation of adaptive ray tracing for radiation
transport \citep{refs}.  To resolve minihalos with at least 100 dark
matter (DM) particles and follow the formation of the first generation
of dwarf galaxies, we use a simulation box of 1 Mpc that has a
resolution of $256^3$.  This gives us a DM mass resolution of 1840
\Ms.  We refine the grid on baryon overdensities of $3 \times
2^{-0.2l}$, where $l$ is the AMR level, resulting in a
super-Lagrangian behavior.  We also refine on a DM overdensity of $3
\times 2^l$ and always resolve the local Jeans length by at least 4
cells.  The latter criterion avoids any artifical fragmentation during
gaseous collapses \citep{Truelove97}.  We initialize the simulation
with \textsl{grafic} \citep{Bertschinger01} at $z = ???$ and use the
cosmological parameters from the 7-year WMAP data \citep{ref}:
$\Omega_M = 0.266$, $\Omega_\Lambda = 0.734$, $\Omega_b = 0.0449$, $h
= 0.71$, $\sigma_8 = 0.81$, and $n = 0.963$ with the variables having
their usual definitions.

We peform two simulations with the same initial conditions.  They
differ only in that one considers primordial cooling, and the other
considers additional cooling in metal-enriched gas.  For primordial
cooling, we use a non-equilbrium chemistry solver with 9 species of
hydrogen, helium, and molecular hydrogen.  For metal cooling, we use a
cooling curve that is calculated from CLOUDY \citep{ref} and using the
method of \citet{ref}.  It is tabulated based on density, temperature,
electron fraction, and metallicity.

\subsection{Star formation}

Simulations have shown that the characteristic mass of Pop III stars
$M_{\rm char} \sim 100~\Ms$.  They form in molecular clouds that
coexist with the dark matter halo center with limited fragmentation
occurring during their collapse; however \citet{Turk10} has recently
shown that Pop III binaries may form in a fraction of such halos.
Furthermore, detailed one-dimensional calculations have shown that
their IMF should follow a Kroupa-like IMF that has a Salpeter slope at
the high-mass end and an exponential cut-off below $M_{\rm char}$
\citep{refs}.

For Pop III stars, we use the same star formation model as
\citet{Abel07} and \citet{Wise08_Gal} but now the stellar mass is
randomly sampled from an IMF with a functional form of
%
\begin{equation}
\label{eqn:imf}
f(M)dM = ...
\end{equation}
%
For reproducibility, we initialize both simulations with the same
random seed and record the number of times the random number generator
\citep[Mersenne twister;][]{ref} has been called.

\subsection{Stellar feedback}
